Argennon: A Scalable Smart Contract Platform

### 2021-06-30 tex/A.tex
Add changes needed by the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

    \chapter{Networking Layer}\label{ch:networking}
    Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
    client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
    crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.
    \note{not yet written...}

@@

    \chapter{Governance}\label{ch:governance}
    The Argennon Decentralized Autonomous Governance system (ADAGs)
    \note{not yet written...}
\end{document}

>>>>

    \chapter{Networking Layer}\label{ch:networking}
    \section{Normal Mode}\label{sec:normal-mode}
    Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
    client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
    crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.
    \note{not yet written...}

@@

    \chapter{Governance}\label{ch:governance}
    \section{ADAGs}\label{sec:adags}
    The Argennon Decentralized Autonomous Governance system (ADAGs)
    \note{not yet written...}
\end{document}

### 2021-06-30 tex/sections/consensus.tex
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
the committee of validators. \textbf{A block is considered final after its next block receives both of
its certificates.}

@@

When an anomaly is detected in the consensus mechanism, the \emph{recovery} protocol is initiated by nodes. The
recovery protocol is designed to be resilient to many types of attacks in order to be able to restore the normal
functionality of the system.

@@

Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
very fast. However, a block is not completely final if it does not have the certificate of the validators.
A certified block by the delegates will not be accepted by the network, if the last block certified by
the validators is behind it more than a certain number of blocks.

@@

Every \texttt{committeeLifeTime} blocks, randomly $m$ committees are selected from
validators, in a way that the total stake of members of each committee is approximately equal, and every
account is a member of \textbf{at least} one committee.

@@

A block certificate issued by some members of a validators' committee is considered valid, if according to
the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
the stake values based on the previous block a malicious committee can affect the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of certificate signers is higher than 80 percent of the total stake
    of the committee members that have \texttt{online} status.
\end{itemize}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part which contains blocks that have a certificate from the committee of delegates but have not received
a certificate from the validators yet.

As we mentioned before, the block with height $n$ needs a certificate from the committee of
validators with index $n$ modulo $m$. To decide about signing a block certificate, a validator
checks the conditional validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
if the block is valid he issues
an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
certificate \textbf{only after} he sees the certificate of the validators of the previous block.

Some validators may also require seeing a certificate from
some community driven committee for the previous block, before they broadcast an "accept" certificate for a block.

So in Argennon, the block validation by committees is performed in parallel, and validators
do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
other hand, the block certificates are published and broadcast sequentially. This ensures that an invalid
fork made by malicious delegates will not receive any certificates from validators.

@@

If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
of the blockchain to be a special \emph{status} block. A status block is a special block which can only contain
status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
the validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
be very low, even zero. Therefore, the status block should not be certified only by online stake.} After applying
the transactions of the status block, the total online stake of validators
must go higher than \texttt{minOnlineStake} threshold.

@@

The recovery protocol is a resilient protocol intended for recovering the Argennon blockchain from critical situations.
In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
during the lifetime of the Argennon blockchain.

\subsection{Estimating A User's Stake}\label{subsec:user's-stake}

>>>>

Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
the committee of validators. \textbf{A block is considered final after it receives both of
its certificates.} In Argennon as long as more than half of the total stake of validators is controlled by honest users,
the probability of discarding a final block is near zero even if all the delegates are malicious.

@@

When an anomaly is detected in the consensus mechanism, the \emph{recovery} protocol is initiated by validators. The
recovery protocol is designed to be resilient to many types of attacks in order to be able to restore the normal
functionality of the system.

@@

Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
very fast. However, a block is not completely final if it does not have the certificate of its validators.
A certified block by the delegates will not be accepted by the network, if the last block certified by
the validators is behind it more than a certain number of blocks.

@@

Every \texttt{committeeLifeTime} number of blocks, randomly $m$ committees are selected from
validators, in a way that the total stake of members of each committee is approximately equal, and every
account is a member of \textbf{at least} one committee.

@@

A block certificate issued by some members of a validators' committee is considered valid, if according to
the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
the stake values based on the previous block a malicious committee can select the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of certificate signers is higher than $3/4$ of the total stake
    of the committee members that have \texttt{online} status.
\end{itemize}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not received
a certificate from the validators yet.

As we mentioned before, the block with height $n$ needs a certificate from the committee of
validators with index $n$ modulo $m$. To decide about signing a block certificate, a validator
checks the conditional validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
if the block is valid he issues
an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
certificate \textbf{only after} he sees the certificate of the validators of the previous block.
Some validators may also require seeing a certificate from
some community driven committee. An honest validator never signs a certificate for two different blocks with the
same height.

So in Argennon, the block validation by committees is performed in parallel, and validators
do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
certificate if the certificate of the previous block has not been published yet. This ensures that an invalid
fork made by malicious delegates will not receive any certificates from validators.

@@

If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
be very low, even zero. Therefore, the status block should not be certified only by online stake.} After applying
the transactions of the status block, the total online stake of validators
must go higher than \texttt{minOnlineStake} threshold.

@@

The recovery protocol is a resilient protocol designed for recovering the Argennon blockchain from critical situations.
In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
during the lifetime of the Argennon blockchain.

The recovery protocol can recover the functionality
of the Argennon blockchain as long as more than 2/3 of the total stake of the system is controlled by honest users
and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
procedures to recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
agreement} protocol.

\subsubsection{Emergency Forking}

The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
from the validators.
This fork needs to be confirmed by the validators and can not discard any blocks that has been already certified
by validators.
A valid fork request is an unexpired request signed by more than half of the
total \texttt{online} stake of the validators.

For forking at block $b$, the reserve committee of delegates
makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $b$.
The height of the fork block is
$b + 1$ so it needs a valid certificate from the committee of validator with the index $b+1$ modulo $m$.

Validators always choose a valid fork block over a block of the main chain. However, as we mentioned before,
a validator never signs a certificate for two different blocks with the same height. Consequently, if a validator
has already signed the block $b$ of the main chain, he will not sign the fork block and vice versa.

When the fork block is broadcast, it is possible that the validators of the committee with index $b+1$ modulo $m$
get divided between the fork block and the block $b+1$ of the main chain, in a way that no block gets enough validators.
This will cause the blockchain to halt. To prevent this, the protocol allows the reserve delegates to revoke a
fork block. After a fork block
is revoked, the validators who voted for it are allowed to vote for another block with the same
height.

To revoke a fork block with height $b+1$, the delegates need to seal the fork by adding a special
\emph{seal block} after the fork block.\footnote{The parent of the seal block must be the fork block.} The seal block
has the height $b+2$ and needs to be certified by the validators' committee with the
index $b+2$ modulo $m$. The fork block is considered revoked only after the seal block is certified by the validators.

A member of the appropriate committee will sign the certificate of a seal block only if he has not already signed
a block after the fork block (i.e., a block with the same parent as the seal block).

The seal block is not a normal block and validators who signed a certificate for a seal block are allowed
to sign a certificate for a block with the same height if the parent of that block is not the fork
block. In other words, the signers of the seal block are not allowed to sign a certificate for
any blocks on the sealed chain, but are allowed to sign certificates for other chains.

As long as more than half of the total online stake of every committee of validators is controlled
by honest users, a malicious committee of delegates can not use the emergency forking procedure to discard blocks
that have a validators' certificate.

Moreover, an honest committee of delegates will always try to perform the emergency forking in such a way that
valid blocks do not get discarded, including blocks that are not certified by the validators yet.

\subsubsection{Emergency Agreement}

The emergency agreement protocol is a resilient protocol for deciding between a set of proposals when no committee of
delegates can be trusted. For initiating the protocol, a validator signs a message containing the subject
of the agreement and a start time.

A validator enters the agreement protocol if he
receives a request that is signed by more than half of the total stake
of the validators and its start time has not passed. The validator calculates
the stake values based on the staking database of the last final block in his blockchain without
considering the \texttt{online}/\texttt{offline} status of validators.

The agreement protocol consists of two phases: the \emph{voting phase}, which selects a single proposal
and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current session ends and a new
agreement session starts. All votes and messages are tagged in a way that the messages of one session can
not be used in another session.

Users cast three type of votes: \emph{$i$-votes}, which are votes that are valid only in
round $i$, \emph{final-votes}, which are votes that are valid in any round, and \emph{c-votes}, which are votes
used only in the confirmation phase.

A user executes the following procedure
in round $r$ of the voting phase:
\paragraph{Voting Phase:}
\begin{itemize}
    \item if the user has not yet final-voted any value, he $r$-votes a single desired proposal.
    \item if he sees more than 2/3 $r$-votes for a proposal $p$, he final-votes $p$.
    \item if he sees more than 2/3 final-votes for a proposal, he goes to the confirmation phase for $p$.
    \item if $clock > r \cdot \lambda$ and $r < k$ user goes to the round $r + 1$ and if $r = k$ user starts a
    new agreement session.
\end{itemize}
\paragraph{Confirmation Phase for Proposal $p$:}
\begin{itemize}
    \item user c-votes $p$.
    \item if he sees more than 2/3 c-votes for $p$ he selects $p$ and ends the agreement protocol.
    \item if $clock > k \cdot \lambda$ user starts a new agreement session.
\end{itemize}

We assume that users have clocks with the same speed, and $\lambda \gg \epsilon$, where $\epsilon$ is the maximum
clock difference between users. We also assume that more than 2/3 of the total stake of the system is controlled
by honest users, and network partitions are resolved after a finite amount of time. With these assumptions it can be
shown that the emergency recovery protocol has the following important properties:
\begin{itemize}
    \item no two users will end the agreement protocol with two different proposals as the result of the agreement.
    \item if honest users can agree upon some proposal value, the agreement protocol will converge to that value
    after a finite number of sessions.
\end{itemize}

When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
confirmation phase could be skipped. In that case, the confirmation of the fork block by the committee of validators
acts like the confirmation phase.

\subsubsection{Initiating the Recovery Protocol}

When a validator does not receive any blocks for \texttt{blockTimeOut} amount of time, or when he sees an
evidence which proves the delegates are malicious, he initiates the recovery protocol.

To do so, first the validator activates the censorship resilient mode of his networking module, then he checks the
validity of the blocks in his blockchain that do not have a validators' certificate and determines the
last valid block of his blockchain.

In the next step, he will sign and broadcast an \textbf{emergency fork request} message, alongside some useful metadata
such as the last valid block of his blockchain and the evidence of delegates' misbehaviour.\footnote{this metadata is
not a part of the fork request.}

If the reserve committee of delegates is already active, or if a validator sees a valid fork request signed
by more than half of the total online stake of the validators, but does not receive the fork
block after a certain amount of time, he will sign and broadcast a request for
\textbf{emergency agreement} on a new reserve committee. The agreement on new delegates usually needs user
interaction and is not a fully automatic process.

The evidence which proves a committee of delegates is malicious is an invalid block that is signed by at least
one delegate:
\begin{itemize}
    \item a block that is not conditionally valid
    \item two different blocks at the same height
    \item a block that has an invalid format
\end{itemize}

\subsection{Estimating Stake Values}\label{subsec:user's-stake}

### 2021-06-30 tex/sections/draft.tex
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

>>>>

If a block receives a "reject" certificate from its committee or does not receive a certificate
after \texttt{certificateTimeOut} period of time, it will be conditionally validated by \textbf{all} validators. If
it does not receive an "accept" certificate, the validators initiate the recovery protocol.

A valid fork block
A validator issues
an "accept" certificate for a fork block at position $b + 1$ (which forks the blockchain at block $b$), if:
\begin{itemize}
    \item the fork block is signed by the reserve committee of delegates and block $b$ is its parent.
    \item the block $b$ has a valid certificate from validators.
    \item the fork block contains a valid fork request according to the staking database of block $b$.
    \item the validator has not signed a certificate for block $b + 1$.
\end{itemize}

will force the fork, in the sense that any validator after receiving it considers the forked chain as the new
valid chain.
Emergency forking may discard at max one block with a valid validator's certificate
from the blockchain, in the worst case scenario.
